Popularity, Mixed Matchings, and Self-duality

### 2017-03-28 CO750-ApproximationAlgorithmsTopics/assignments/presentation.tex
last minute fixes

\begin{frame}
\frametitle{Fractional Matchings}
We can extend this to the $\text{\alert{fractional matching}}$ case:
$$FM_{G} = \{x \in \R^m_+ : x(\delta(u)) \leq 1, \forall u \in V(G)\}.$$
We extend the voting function by defining for all $\alert{x \in FM_G}$, for all $\alert{u,v' \in V(G)}$:
$$vote_u(x,v) = \sum_{v\in V(G)} x_{uv} vote_u(v,v').$$ 
We extend the vote aggregation defining for all $\alert{x, y \in FM_{G}}$:
$$\Delta(x,y) = \sum_{u \in V(G)} \sum_{uv, uv' \in \delta(u)} x_{uv}y_{uv'} vote_u(v,v').$$
The polytope of all $\text{\alert{fractional popular matchings}}$ is 
$$P_G := \{x \in FM_{G}: \Delta(x,y) \geq 0, \forall y \in FM_G\}.$$
\end{frame}

\begin{frame}
\frametitle{Mixed Popular Matching Example}
In $\text{\alert{bipartite}}$ graphs, each $\alert{x \in FM_G}$ is a convex combination of matchings. But observe that each $\alert{x\in P_G}$ need not be a convex combination of $\text{\alert{popular}}$ matchings.
\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
\begin{figure}
\begin{tikzpicture} 
\node[] (a0) at (0,5) {$a_0$};
    \node[] (a1) at (0,4) {$a_1$};
    \node[] (a2) at (0,3) {$a_2$};

@@

\begin{frame}
\frametitle{Extended Formulation}
Fix a popular fractional matching $\alert{y}$. Then $\alert{\Delta(x,y)}$ is a linear function in $\alert{x}$. Since $\alert{y}$ is popular $\alert{\Delta(x,y) \leq 0}$ with equality when $\alert{x}$ is popular.
\begin{columns}[T]
\begin{column}{.48\textwidth}
\begin{align*}
\max &\Delta(x,y) \\
\text{s.t.} \sum_{e \in \delta(u)} x_e &= 1 &\forall u \in V(G) \\
x_e &\geq 0 &\forall e \in E
\end{align*}
\end{column}
\begin{column}{.48\textwidth}
\begin{align*}
\min &\sum_{u \in V(G)} \alpha_u  \\
\text{s.t.} \alpha_a + \alpha_b &\geq vote_a(b,y) \\&+ vote_b(a,y) &\forall ab \in E(G) \\
\alpha_u &\geq vote_u(\ell(u), y) &\forall u \in V(G).
\end{align*}
\end{column}
\end{columns}
Where $\alert{\ell(u)}$ is a last choice vertex for each $\alert{u}$ so that each vertex is fully matched in any fractional matching.

@@

\begin{frame}
\frametitle{Self-Duality}
It is straightforward to see that $\alert{(LP)}$ is self-dual (exercise). This implies, by $\text{\alert{complementary slackness}}$, that the covering constraint for $\alert{ab}$ is tight when $\alert{x_{ab} > 0}$ for any popular fractional matching $\alert{x}$.

@@

So the hypothesis says each $\text{\alert{popular matching}}$ is perfect and, in the fractional setting, for each $\alert{u\in A \cup B}$, $\alert{x_{u\ell(u)} = 0}$.

@@

\begin{frame}
\frametitle{Sweeping Procedure}
We write $\alert{x}$ as a convex combination of $\text{\alert{popular matchings}}$:
\begin{enumerate}
\item Initialize $\alert{i=0}$.
\item Sweep a vertical line from left to right across $\alert{T}$. Denote its distance by $\alert{t}$. 
\item Whenever a new cell is encountered fix matching $\alert{M^i = M_t}$ and fix $\alert{t_i = t}$. 
\end{enumerate}
Say we find $\alert{k}$ such matchings. We can then write $\alert{x}$ as:
$$x = t_1\chi(M^0) + (t_2 - t_1)\chi(M^1) + \dots + (1 - t_{k-1}) \chi(M^{k-1}).$$
This constructive procedure runs in $\alert{O(|E| + |V|)}$ time.
\end{frame}
\section{Half-Integrality of $P_G$ in a general instance}

@@

Is there an algorithm for deciding if a general graph admits a $\text{\alert{popular matching}}$ under $\text{\alert{strict}}$ preferences?
\end{frame}

>>>>

\begin{frame}
\frametitle{Fractional Matchings}
We can extend this to the $\text{\alert{fractional matching}}$ case:
$$FM_{G} = \{x \in \R^m_+ : x(\delta(u)) \leq 1, \forall u \in V(G)\}.$$
We extend the voting function by defining for all $\alert{x \in FM_G}$, for all $\alert{u,v' \in V(G)}$:
$$vote_u(x,v) = \sum_{uv'\in \delta(u)} x_{uv'} vote_u(v',v) \text{ and } vote_u(v,x) = -vote_u(x,v).$$ 
We extend the vote aggregation defining for all $\alert{x, y \in FM_{G}}$:
$$\Delta(x,y) = \sum_{u \in V(G)} \sum_{uv \in \delta(u)} x_{uv}vote_u(v,y).$$
The polytope of all $\text{\alert{fractional popular matchings}}$ is 
$$P_G := \{x \in FM_{G}: \Delta(x,y) \geq 0, \forall y \in FM_G\}.$$
\end{frame}

\begin{frame}
\frametitle{Mixed Popular Matching Example}
In $\text{\alert{bipartite}}$ graphs, each $\alert{x \in FM_G}$ is a convex combination of matchings. But each $\alert{x\in P_G}$ need not be a convex combination of $\text{\alert{popular}}$ matchings.
\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
\begin{figure}
\begin{tikzpicture} 
\node[] (a0) at (0,5) {$a_0$};
    \node[] (a1) at (0,4) {$a_1$};
    \node[] (a2) at (0,3) {$a_2$};

@@

\begin{frame}
\frametitle{Extended Formulation}
Fix a popular fractional matching $\alert{y}$. Then $\alert{\Delta(x,y)}$ is a linear function in $\alert{x}$. Since $\alert{y}$ is popular $\alert{\Delta(x,y) \leq 0}$ with equality when $\alert{x}$ is popular.
\begin{columns}[T]
\begin{column}{.48\textwidth}
\begin{align*}
\max &\Delta(x,y) \\
\text{s.t.} \sum_{e \in \delta(u)} x_e &= 1 &\forall u \in V(G) \\
x_e &\geq 0 &\forall e \in E
\end{align*}
\end{column}
\begin{column}{.48\textwidth}
\begin{align*}
\min &\sum_{u \in V(G)} \alpha_u  \\
\text{s.t.} \alpha_a + \alpha_b &\geq vote_a(b,y) \\&+ vote_b(a,y) &\forall ab \in E(G) \\
\alpha_u &\geq vote_u(\ell(u), y) &\forall u \in V(G).
\end{align*}
\end{column}
\end{columns}
Where $\alert{\ell(u)}$ is an appended last choice vertex for each $\alert{u}$.

@@

\begin{frame}
\frametitle{Self-Duality}
It is straightforward to see that $\alert{(LP)}$ is self-dual (exercise). By $\text{\alert{complementary slackness}}$: the covering constraint for $\alert{ab}$ is tight when $\alert{x_{ab} > 0}$ for any popular fractional matching $\alert{x}$.

@@

Hypothesis says each $\text{\alert{popular matching}}$ is perfect and, in fractional setting, for each $\alert{u\in A \cup B}$, $\alert{x_{u\ell(u)} = 0}$.

@@

\begin{frame}
\frametitle{Sweeping Procedure}
We write $\alert{x}$ as a convex combination of $\text{\alert{popular matchings}}$:
\begin{enumerate}
\item Initialize $\alert{i=0}$.
\item Sweep a vertical line from left to right across $\alert{T}$. Denote its distance by $\alert{t}$. 
\item Whenever a new cell is encountered fix matching $\alert{M^i = M_t}$ and fix $\alert{t_i = t}$. 
\end{enumerate}
Say we find $\alert{k}$ such matchings. We can then write $\alert{x}$ as:
$$x = t_1\chi(M^0) + (t_2 - t_1)\chi(M^1) + \dots + (1 - t_{k-1}) \chi(M^{k-1}).$$
\end{frame}
\section{Half-Integrality of $P_G$ in a general instance}

@@

Is there an algorithm for deciding if a general graph admits a $\text{\alert{popular matching}}$ under $\text{\alert{strict}}$ preferences?

$$ $$
Can we say anything better for $\text{\alert{roommates}}$ instances with a $\text{\alert{perfect stable matching}}$?
\end{frame}
